Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference

### 2025-01-09 src/tex/ms.tex
[showyourwork] overleaf sync

\begin{figure}[]
     \centering
     \includegraphics[width=\columnwidth]{figures/posterior_ape.png}
     \vspace{-.5cm}
     \caption{Absolute percentage error of the neural posterior density estimate for a single star. Different coloured histograms show the full error distribution for all 6 parameters of interest with the median values highlighted by the vertical dashed lines. The box plots show again the first and third quartiles with the median represented by a vertical line and the whiskers extending from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box. The global parameters of main interest for this work are shown by the light blue and red histogram.}
     \label{fig:ape_NN}
     \script{train_sbi.py}
\end{figure}

For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate with 1000 data points (see Fig.~\ref{fig:posterior_samples} in the appendix). We then compare the posterior mean to the ground truth value. For a single star observation, our NPE has an APE of \variable{output/posterior_APE.txt} when looking at all 6 parameters of interest. If we restrict ourselves to only the global two parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ our NPE achieves an APE of \variable{output/global_posterior_APE_NPE_C.txt} as we show in Fig.~\ref{fig:posterior_APE}. Looking at Fig.~\ref{fig:posterior_APE} we find that for an individual star the accuracy of the NPE is not particularly high. However, for the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ we can boost the accuracy by combining the inference for many stars of the same galaxy.

>>>>

\begin{figure}[]
     \centering
     \includegraphics[width=\columnwidth]{figures/ape_posterior_NPE_C.pdf}
     \vspace{-.5cm}
     \caption{Absolute percentage error of the neural posterior density estimate for a single star. Different coloured histograms show the full error distribution for all 6 parameters of interest with the median values highlighted by the vertical dashed lines. The box plots show again the first and third quartiles with the median represented by a vertical line and the whiskers extending from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box. The global parameters of main interest for this work are shown by the light blue and red histogram.}
     \label{fig:ape_NN}
     \script{evaluate_posterior.py}
\end{figure}

For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate with 1000 data points (see Fig.~\ref{fig:posterior_samples} in the appendix). We then compare the posterior mean to the ground truth value. For a single star observation, our NPE has an APE of \variable{output/posterior_APE.txt} when looking at all 6 parameters of interest. If we restrict ourselves to only the global two parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ our NPE achieves an APE of \variable{output/global_posterior_APE.txt} as we show in Fig.~\ref{fig:posterior_APE}. Looking at Fig.~\ref{fig:posterior_APE} we find that for an individual star the accuracy of the NPE is not particularly high. However, for the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ we can boost the accuracy by combining the inference for many stars of the same galaxy.
